---
title: "Analysis notebook"
subtitle: "Use this to add your intermediate data analysis, plots, etc."
author: "Jinwook Bae, Miriam Haf, Xuchen Shen, Andreas Sünder"
lang: en
format: html
---

```{r setup, include=FALSE}
source("source_all.R")
download_data()

COUNTRIES_FOR_COMPLETION <- c("Austria", "Vietnam", "Tunisia", "Brazil")
YEARS_FOR_COMPLETION <- c(2010, 2015, 2020)

world <- ne_countries(scale = "medium", returnclass = "sf")
```

# Analysis of generic inequality data

```{r}
gii <- load_owid_gii_data()

gii.latest <- gii[year == 2023]
gii_world <- merge(world, gii.latest, by.x = "iso_a3", by.y = "countryCode", all.x = TRUE)

ggplot(gii_world) +
  geom_sf(aes(fill = value), color = "grey30", linewidth = 0.1) +
  scale_fill_distiller(
    palette = "YlOrRd", direction = 1,
    name = "GII", na.value = "grey80",
    limits = c(0, 1)
  ) +
  coord_sf(ylim = c(-56, 90)) +
  labs(title = "Gender Inequality Index, 2023") +
  theme_light() +
  theme(axis.text = element_blank(), axis.ticks = element_blank())
```

# Analysis of education data

## Completion rates (observed data)

```{r}
completion <- load_uis_completion_rates()

# Global completion rates, averaged across available countries
completion.global <- completion[year == 2021][, .(value = mean(value)), by = .(educationLevel, sex)]
ggplot(completion.global, aes(x = educationLevel, y = value, fill = sex)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(title = "Global completion rates by gender, 2021", x = "education level", y = "average completion rate") +
  theme_light()

# Completion rates for selected countries
completion.countries <- completion[country %in% COUNTRIES_FOR_COMPLETION & year == 2021]
ggplot(completion.countries, aes(x = country, y = value, fill = sex)) +
  geom_col(position = "dodge") +
  facet_wrap(~educationLevel, ncol = 1) +
  labs(title = "Completion rates for selected countries, 2022", y = "completion rate") +
  theme_light()

# over time (averaged over all countries)
completion.time <- completion[year %in% YEARS_FOR_COMPLETION,
  .(educationLevel, country, sex, year, value, n = uniqueN(interaction(educationLevel, sex, year))),
  by = countryCode
][n == 3 * 2 * 3][, n := NULL][,
  .(value = mean(value)),
  by = .(educationLevel, sex, year)
]
ggplot(completion.time, aes(x = year, y = value, color = sex, linetype = educationLevel)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = YEARS_FOR_COMPLETION) +
  scale_y_continuous(breaks = seq(65, 100, by = 5)) +
  labs(title = "Average completion rates over time", subtitle = "Derived from selected countries with data availability across all years", y = "average completion rate") +
  theme_light()
```

## Gender parity index

```{r}
gpi <- load_owid_gpi_data()

# we outsource the repeating tasks for each education level
# to avoid duplicated code segments
plot_gpi_map <- function(data, level) {
  gpi_filtered <- data[year == 2020 & educationLevel == level]
  world_merged <- merge(world, gpi_filtered, by.x = "iso_a3", by.y = "countryCode", all.x = TRUE)

  ggplot(world_merged) +
    geom_sf(aes(fill = value), color = "grey30", linewidth = 0.1) +
    scale_fill_gradient2(
      midpoint = 1, low = "#b2182b", mid = "#ffffbf", high = "#2166ac",
      name = "GPI", na.value = "grey80",
      limits = c(0.5, 1.5), oob = scales::squish
    ) +
    coord_sf(ylim = c(-56, 90)) +
    labs(title = paste("Gender Parity Index -", level, "education"), subtitle = "Expressed as the ratio between the GER of females and males") +
    theme_light() +
    theme(legend.position = "bottom", legend.key.width = unit(2, "cm"), legend.title = element_text(vjust = 0.75), axis.text = element_blank(), axis.ticks = element_blank())
}

plot_gpi_map(gpi, "Primary")
plot_gpi_map(gpi, "Lower secondary")
plot_gpi_map(gpi, "Upper secondary")
plot_gpi_map(gpi, "Tertiary")

# TODO: Visualize gender parity index over time and/or across selected countries
```

# Analysis of Care Responsbility Share data

We know utilize a ILOSTAT dataset that has more data points to analyze the
difference of care responsibilities in specific countries and along the
time line. We utilize facet wrapping of line plot portraying values given in data.

```{r}
care_responsibilities_share <- load_care_resp_share_data()

# We first interpret which country has most values.
care_responsibilities_share[sex.label %in% c("Male", "Female") & !is.na(obs_value),
   .(n_years = uniqueN(time)),
   by = ref_area.label
][order(-n_years)] |> head(30)

# facet wrapped graphic
plot_care_share_facet(care_responsibilities_share)

# Let us analyze the gap in care responsibilities between genders for
# extreme cases. Which countries show clear inequality in share towards women
# and which don't ?

plot_care_gap_ranking(care_responsibilities_share)

# Now lets look at the gap amongst european countries (discuss if we should include)
care_responsibilities_share[sex.label %in% c("Male", "Female") & !is.na(obs_value),
   .N,
   by = ref_area.label
][order(-N)] |> head(50)

# use countries with long time series
plot_care_share_trend(care_responsibilities_share, "South Africa")
plot_care_share_trend(care_responsibilities_share, "Chile")
plot_care_share_trend(care_responsibilities_share, "Australia")
plot_care_share_trend(care_responsibilities_share, "Canada")

plot_care_share_gii_correlation(care_responsibilities_share)
```

# Analysis of Domestic Work Time data

In this task, I utilize the OWID data for percentage of day used on unpaid and domestic work.
Weights are not used as we are dealing with sparse data where not all countries at all times
are covered. We simply want to identify if there is a differnce in amount of care work between
male and female. We take the mean value across all years for each country and then the median
from the means in each continent and plot a bar graph to identify the differnce.
A more numerical analysis is done within the gt table with gaps and ratios.

```{r}
owid_domestic_work_time <- load_domestic_work_time_data()

plot_domestic_work_region(owid_domestic_work_time)
table_domestic_work_region(owid_domestic_work_time)
```

# Analysis of Labour Force Participation data

```{r}
labour_force_with_educ <- load_labour_force_with_educ_data()

#First data cleaning for labour force participation rate
labour_force_with_educ_cleaned <- prepare_lfpr_data(labour_force_with_educ)

#Research on LFPR for women in year 2024
labour_force_with_educ_cleaned %>% filter(sex == "Female") %>%
  summarize(Median = median(rate), Mean =  mean(rate), Min = min(rate),
            Max = max(rate))
labour_force_with_educ_cleaned %>% filter(sex == "Female",
                                  when_any(rate == 14.106, rate == 81.691))


#Preparing data for first analysis
lf_total_2024_comparison <-
  prepare_lfpr_total_2024_comp(labour_force_with_educ_cleaned)

#Graph: LFPR Male vs. Female in 2024 in five countries: GER, AUS, USA, IRN, TZA
lf_total_2024_comparison %>% 
  ggplot(aes(x = country, y = rate, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Labour Force Participation Rate in 2024 by Gender",
        x = "Country",
        y = "Labour Force Participation Rate in %",
        fill = "Biological Gender") +
  scale_y_continuous(limits = c(0,90), breaks = seq(0,90,20))

#Preparing data for second analysis
lf_GER_2010_to_2024 <- 
  prepare_lfpr_GER_2010_to_2024(labour_force_with_educ_cleaned)

#Graph: LFPR Male vs.Female in GER from 2010 until 2024
lf_GER_2010_to_2024 %>%
  ggplot(aes(x = year, y = rate, colour = sex)) + geom_line() +
  labs(title = "Labour Force Participation Rate in Germany between 2010 and 2024",
       x = "Year",
       y = "Labour Force Participation Rate in %",
       colour = "Biological Gender") +
  scale_y_continuous(limits = c(0,70), breaks = seq(0,70,10))
```

