---
title: "Analysis notebook"
subtitle: "Use this to add your intermediate data analysis, plots, etc."
author: "Jinwook Bae, Miriam Haf, Xuchen Shen, Andreas Sünder"
lang: en
format: html
---

```{r setup, include=FALSE}
source("source_all.R")
download_data()

COUNTRIES_FOR_COMPLETION <- c("Yemen", "Afghanistan", "Sweden", "Brazil")
COUNTRIES_FOR_COMPLETION_LATEST <- c("Austria", "Vietnam", "Tunisia", "Brazil")
YEARS_FOR_COMPLETION <- c(2010, 2015, 2020)

world <- ne_countries(scale = "medium", returnclass = "sf")
```

# Analysis of generic inequality data

```{r}
gii <- load_owid_gii_data()

gii.latest <- gii[year == 2023]
gii_world <- merge(world, gii.latest, by.x = "iso_a3", by.y = "countryCode", all.x = TRUE)

ggplot(gii_world) +
  geom_sf(aes(fill = value), color = "grey30", linewidth = 0.1) +
  scale_fill_distiller(
    palette = "YlOrRd", direction = 1,
    name = "GII", na.value = "grey80",
    limits = c(0, 1)
  ) +
  coord_sf(ylim = c(-56, 90)) +
  labs(title = "Gender Inequality Index, 2023") +
  theme_light() +
  theme(axis.text = element_blank(), axis.ticks = element_blank())
```

# Analysis of education data

## Completion rates (observed data)

```{r}
completion <- load_uis_completion_rates()

# Global completion rates, averaged across available countries
completion.global.latest <- completion[year == 2021][, .(value = mean(value)), by = .(educationLevel, sex)]
ggplot(completion.global.latest, aes(x = educationLevel, y = value, fill = sex)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(title = "Average completion rates by sex, 2021", x = "Education level", y = "Average completion rate", fill = "Sex") +
  theme_light()

# Completion rates for selected countries (2021)
# Might now be optional
completion.countries.latest <- completion[country %in% COUNTRIES_FOR_COMPLETION_LATEST & year == 2021]
ggplot(completion.countries.latest, aes(x = country, y = value, fill = sex)) +
  geom_col(position = "dodge") +
  facet_wrap(~educationLevel, ncol = 1) +
  labs(title = "Completion rates for selected countries, 2021", x = "Country", y = "Completion rate", fill = "Sex") +
  theme_light()

# Completion rates for other selected countries (based on what I found in the GII plot)
# Fortunately, for all countries, there is data available for the same years
# across all education levels
# To make things easier, we take the mean value for both genders to and aggregate across
# all education levels
completion.countries <- completion[country %in% COUNTRIES_FOR_COMPLETION, .(value = mean(value)), by = .(country, year, sex)]
ggplot(completion.countries, aes(x = year, y = value, color = sex)) +
  geom_line() +
  geom_point() +
  facet_wrap(~country, nrow = 2) +
  labs(title = "Completion rates for selected countries over time", subtitle = "Averaged across primary, lower secondary, and upper secondary education", x = "Year", y = "Average completion rate", color = "Sex") +
  theme_light()

# Completion rates over time (averaged over all countries)
# the code below makes sure that before averaging, we only select those countries
# that have complete data for 2010, 2015, and 2020 (called a "balanced panel").
# Otherwise, calculating the average based on different populations would be inadmissible.
completion.time <- completion[year %in% YEARS_FOR_COMPLETION,
  .(educationLevel, country, sex, year, value, n = uniqueN(interaction(educationLevel, sex, year))),
  by = countryCode
][n == 3 * 2 * 3][, n := NULL][,
  .(value = mean(value)),
  by = .(educationLevel, sex, year)
]
ggplot(completion.time, aes(x = year, y = value, color = sex, linetype = educationLevel)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = YEARS_FOR_COMPLETION) +
  scale_y_continuous(breaks = seq(65, 100, by = 5)) +
  labs(title = "Completion rates over time", subtitle = "Averaged across countries with complete data for 2010, 2015, and 2020", x = "Year", y = "Average completion rate", color = "Sex", linetype = "Education level") +
  theme_light()
```

## Gender parity index

```{r}
gpi <- load_owid_gpi_data()

# we outsource the repeating tasks for each education level
# to avoid duplicated code segments
plot_gpi_map <- function(data, level) {
  gpi_filtered <- data[year == 2020 & educationLevel == level]
  world_merged <- merge(world, gpi_filtered, by.x = "iso_a3", by.y = "countryCode", all.x = TRUE)

  ggplot(world_merged) +
    geom_sf(aes(fill = value), color = "grey30", linewidth = 0.1) +
    scale_fill_gradient2(
      midpoint = 1, low = "#b2182b", mid = "#ffffbf", high = "#2166ac",
      name = "GPI", na.value = "grey80",
      limits = c(0.5, 1.5), oob = scales::squish
    ) +
    coord_sf(ylim = c(-56, 90)) +
    labs(title = paste("Gender Parity Index -", level, "education"), subtitle = "Ratio of female to male gross enrollment ratio") +
    theme_light() +
    theme(legend.position = "bottom", legend.key.width = unit(2, "cm"), legend.title = element_text(vjust = 0.75), axis.text = element_blank(), axis.ticks = element_blank())
}

plot_gpi_map(gpi, "Primary")
plot_gpi_map(gpi, "Lower secondary")
plot_gpi_map(gpi, "Upper secondary")
plot_gpi_map(gpi, "Tertiary")

# TODO: Visualize gender parity index over time and/or across selected countries (optional)
```

# Analysis of Care Responsbility Share data

We know utilize a ILOSTAT dataset that has more data points to analyze the
difference of care responsibilities in specific countries and along the
time line. We utilize facet wrapping of line plot portraying values given in data.

```{r}
care_responsibilities_share <- load_care_resp_share_data()

# We first interpret which country has most values.
care_responsibilities_share[sex.label %in% c("Male", "Female") & !is.na(obs_value),
  .(n_years = uniqueN(time)),
  by = ref_area.label
][order(-n_years)] |> head(30)

# facet wrapped graphic
plot_care_share_facet(care_responsibilities_share)

# Let us analyze the gap in care responsibilities between genders for
# extreme cases. Which countries show clear inequality in share towards women
# and which don't ?

plot_care_gap_ranking(care_responsibilities_share)

# Now lets look at the gap amongst european countries (discuss if we should include)
care_responsibilities_share[sex.label %in% c("Male", "Female") & !is.na(obs_value),
  .N,
  by = ref_area.label
][order(-N)] |> head(50)

# use countries with long time series
plot_care_share_trend(care_responsibilities_share, "South Africa")
plot_care_share_trend(care_responsibilities_share, "Chile")
plot_care_share_trend(care_responsibilities_share, "Australia")
plot_care_share_trend(care_responsibilities_share, "Canada")

plot_care_share_gii_correlation(care_responsibilities_share)
```

# Analysis of Domestic Work Time data

In this task, I utilize the OWID data for percentage of day used on unpaid and domestic work.
Weights are not used as we are dealing with sparse data where not all countries at all times
are covered. We simply want to identify if there is a differnce in amount of care work between
male and female. We take the mean value across all years for each country and then the median
from the means in each continent and plot a bar graph to identify the differnce.
A more numerical analysis is done within the gt table with gaps and ratios.

```{r}
owid_domestic_work_time <- load_domestic_work_time_data()

plot_domestic_work_region(owid_domestic_work_time)
table_domestic_work_region(owid_domestic_work_time)
```

# Analysis of Labour Force Participation data

```{r}
labour_force_with_educ <- load_labour_force_with_educ_data()

<<<<<<< labor-force-participation
#First data cleaning for labour force participation rate
labour_force_with_educ_cleaned <- prepare_lfpr_data(labour_force_with_educ)

#Research on LFPR for women in year 2024
labour_force_with_educ_cleaned %>% filter(sex == "Female") %>%
  summarize(Median = median(rate), Mean =  mean(rate), Min = min(rate),
            Max = max(rate))
labour_force_with_educ_cleaned %>% filter(sex == "Female",
                                  when_any(rate == 14.106, rate == 81.691))


#Preparing data for first analysis
lf_total_2024_comparison <-
  prepare_lfpr_total_2024_comp(labour_force_with_educ_cleaned)

#Graph: LFPR Male vs. Female in 2024 in five countries: GER, AUS, USA, IRN, TZA
lf_total_2024_comparison %>% 
  ggplot(aes(x = country, y = rate, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Labour Force Participation Rate in 2024 by Gender",
        x = "Country",
        y = "Labour Force Participation Rate in %",
        fill = "Biological Gender") +
  scale_y_continuous(limits = c(0,90), breaks = seq(0,90,20))

#Preparing data for second analysis
lf_GER_2010_to_2024 <- 
  prepare_lfpr_GER_2010_to_2024(labour_force_with_educ_cleaned)

#Graph: LFPR Male vs.Female in GER from 2010 until 2024
lf_GER_2010_to_2024 %>%
  ggplot(aes(x = year, y = rate, colour = sex)) + geom_line() +
  labs(title = "Labour Force Participation Rate in Germany between 2010 and 2024",
       x = "Year",
       y = "Labour Force Participation Rate in %",
       colour = "Biological Gender") +
  scale_y_continuous(limits = c(0,70), breaks = seq(0,70,10))
=======
# Umbenennen relevanter Variablen
labour_force_with_educ <- rename(labour_force_with_educ,
  country = ref_area.label,
  data_source = source.label,
  sex = sex.label,
  education = classif1.label,
  year = time,
  rate = obs_value
)

# Selektieren relevanter Variablen
labour_force_with_educ <- select(
  labour_force_with_educ,
  country, data_source, sex, education, year, rate
)

# Filtern nach einer Skala der Bildung
labour_force_with_educ <- filter(
  labour_force_with_educ,
  str_detect(education, "ISCED-11")
)

# Filtern nach Sex, nur Männer und Frauen
labour_force_with_educ <- filter(
  labour_force_with_educ,
  when_any(sex == "Female", sex == "Male")
)


# Filtern - keine NAs für Labour Force Participation Rate
labour_force_with_educ <- filter(labour_force_with_educ, !is.na(rate))

# Übersicht - Labour Force Participation Rate im Jahr 2024 (Total)
labour_force_total_2024 <- filter(
  labour_force_with_educ,
  year == 2024, str_detect(education, "Total")
)

# Untersuchen der Variable LFPR für Frauen im Jahr 2024
labour_force_total_2024 |>
  filter(sex == "Female") |>
  summarize(
    Median = median(rate), Mean = mean(rate), Min = min(rate),
    Max = max(rate)
  )
labour_force_total_2024 |> filter(
  sex == "Female",
  when_any(rate == 14.106, rate == 81.691)
)

plot_lfpr_countries_2024 <- function(data) {
  data |> ggplot(aes(x = country, y = rate, fill = sex)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(
      title = "Labour Force Participation Rate in 2024 by Gender",
      x = "Country",
      y = "Labour Force Participation Rate in %",
      fill = "Biological Gender"
    ) +
    scale_y_continuous(limits = c(0, 90), breaks = seq(0, 90, 20))
}

# Grafik: LFPR Männer vs. Frauen im Jahr 2024 für Deutschland, USA, Tansania, Iran, Australien
labour_force_total_2024 |>
  filter(when_any(
    country == "Tanzania, United Republic of",
    country == "Germany",
    country == "United States of America",
    country == "Iran (Islamic Republic of)",
    country == "Australia"
  )) |>
  ggplot(aes(x = country, y = rate, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Labour Force Participation Rate in 2024 by Gender",
    x = "Country",
    y = "Labour Force Participation Rate in %",
    fill = "Biological Gender"
  ) +
  scale_y_continuous(limits = c(0, 90), breaks = seq(0, 90, 20))

# Grafik - LFPR Männer vs. Frauen in Deutschland von 2014 bis 2024
labour_force_with_educ |>
  filter(
    str_detect(education, "Total"),
    country == "Germany", year >= 2014
  ) |>
  ggplot(aes(x = year, y = rate, colour = sex)) +
  geom_line() +
  labs(
    title = "Labour force participation rate in Germany, 2014-2024",
    x = "Year",
    y = "Labour force participation rate (%)",
    colour = "Sex"
  ) +
  scale_y_continuous(limits = c(0, 70), breaks = seq(0, 70, 10))
>>>>>>> main
```

