---
title: "Analyse-Notebook"
author: "Jinwook Bae, Miriam Haf, Xuchen Shen, Andreas Sünder"
lang: de
format: html
---

```{r setup, include=FALSE}
source("source_all.R")
download_data()
```

# Analysis of Care Responsbility Share data

We know utilize a ILOSTAT dataset that has more data points to analyze the
difference of care responsibilities in specific countries and along the
time line. We utilize facet wrapping of line plot portraying values given in data.

```{r}
care_responsibilities_share <- load_care_resp_share_data()

# We first interpret which country has most values.
care_responsibilities_share[sex.label %in% c("Male", "Female") & !is.na(obs_value),
   .(n_years = uniqueN(time)),
   by = ref_area.label
][order(-n_years)] |> head(30)

# facet wrapped graphic
plot_care_share_facet(care_responsibilities_share)

# Let us analyze the gap in care responsibilities between genders for
# extreme cases. Which countries show clear inequality in share towards women
# and which don't ?

plot_care_gap_ranking(care_responsibilities_share)

# Now lets look at the gap amongst european countries (discuss if we should include)
care_responsibilities_share[sex.label %in% c("Male", "Female") & !is.na(obs_value),
   .N,
   by = ref_area.label
][order(-N)] |> head(50)

# use countries with long time series
plot_care_share_trend(care_responsibilities_share, "South Africa")
plot_care_share_trend(care_responsibilities_share, "Chile")
plot_care_share_trend(care_responsibilities_share, "Australia")
plot_care_share_trend(care_responsibilities_share, "Canada")

plot_care_share_gii_correlation(care_responsibilities_share)
```

# Analysis of Domestic Work Time data

In this task, I utilize the OWID data for percentage of day used on unpaid and domestic work.
Weights are not used as we are dealing with sparse data where not all countries at all times
are covered. We simply want to identify if there is a differnce in amount of care work between
male and female. We take the mean value across all years for each country and then the median
from the means in each continent and plot a bar graph to identify the differnce.
A more numerical analysis is done within the gt table with gaps and ratios.

```{r}
owid_domestic_work_time <- load_domestic_work_time_data()

plot_domestic_work_region(owid_domestic_work_time)
table_domestic_work_region(owid_domestic_work_time)
```

