---
title: "Gender Inequality"
subtitle: "Statistical project (LMU Munich)"
author: "Jinwook Bae, Miriam Haf, Xuchen Shen, Andreas Sünder"
date: today
date-format: "MMMM D, YYYY"
lang: en
execute:
  echo: false
format:
  revealjs:
    fontsize: 32px
    theme: default
    slide-number: true
    logo: assets/logo.svg
    css: presentation.css
---

```{r setup, include=FALSE}
# Put any `library()` calls (or similar here)
# We can't put those into the setup block below, as the functions loaded by them
# are not restored correctly when cached
source("source_all.R")
download_data()
```

```{r data-prep, include=FALSE, cache=TRUE}
# You can put any constant values here
COUNTRIES_FOR_COMPLETION <- c("Yemen", "Afghanistan", "Sweden", "Brazil")
COUNTRIES_FOR_COMPLETION_LATEST <- c("Austria", "Vietnam", "Tunisia", "Brazil")
YEARS_FOR_COMPLETION <- c(2010, 2015, 2020)

world <- ne_countries(scale = "medium", returnclass = "sf")

gii <- load_owid_gii_data()
gii.latest <- gii[year == 2023]
gii_world <- merge(world, gii.latest, by.x = "iso_a3", by.y = "countryCode", all.x = TRUE)

completion <- load_uis_completion_rates()
completion.global.latest <- completion[year == 2021][, .(value = mean(value)), by = .(educationLevel, sex)]
completion.countries <- completion[country %in% COUNTRIES_FOR_COMPLETION, .(value = mean(value)), by = .(country, year, sex)]
completion.countries.latest <- completion[country %in% COUNTRIES_FOR_COMPLETION_LATEST & year == 2021]

gpi <- load_owid_gpi_data()
# we outsource the repeating tasks for each education level
# to avoid duplicated code segments
plot_gpi_map <- function(data, level) {
  gpi_filtered <- data[year == 2020 & educationLevel == level]
  world_merged <- merge(world, gpi_filtered, by.x = "iso_a3", by.y = "countryCode", all.x = TRUE)

  ggplot(world_merged) +
    geom_sf(aes(fill = value), color = "grey30", linewidth = 0.1) +
    scale_fill_gradient2(
      midpoint = 1, low = "#b2182b", mid = "#ffffbf", high = "#2166ac",
      name = "GPI", na.value = "grey80",
      limits = c(0.5, 1.5), oob = scales::squish
    ) +
    coord_sf(ylim = c(-56, 90)) +
    labs(title = paste("Gender Parity Index -", level, "education"), subtitle = "Expressed as the ratio between the GER of females and males") +
    theme(legend.title = element_text(vjust = 0.75), axis.text = element_blank(), axis.ticks = element_blank())
}

# this is an experimental function that is used by
# the combined GPI plot as it requires inducing some custom
# spacing to make the map use the available space efficiently
gpi_subplot <- function(level) {
  plot_gpi_map(gpi, level) +
    labs(title = level, subtitle = NULL) +
    theme(
      plot.title = element_text(size = 12, hjust = 0.5, margin = margin(b = 3)),
      plot.margin = margin(t = 8, r = 2, b = 2, l = 2)
    )
}

labour_force_with_educ <- load_labour_force_with_educ_data()
labour_force_with_educ_cleaned <- prepare_lfpr_data(labour_force_with_educ)

lf_total_2024_comparison <-
  prepare_lfpr_total_2024_comp(labour_force_with_educ_cleaned)

lf_GER_2010_to_2024 <-
  prepare_lfpr_GER_2010_to_2024(labour_force_with_educ_cleaned)

lf_2024_by_education <-
  prepare_lfpr_2024_by_education(labour_force_with_educ_cleaned)

wage_gap <- load_wage_gap_data()
trend_gap <- prepare_wage_gap_trend(wage_gap)

latest_year <- max(wage_gap$time)
top_countries <- prepare_wage_gap_top_countries(wage_gap, latest_year)

first_year <- min(wage_gap$time)
last_year <- max(wage_gap$time)
change_full <- prepare_wage_gap_change(wage_gap, first_year, last_year)

change_1920 <- prepare_wage_gap_change(wage_gap, 2019, 2020)
```

# Motivation

---

```{r gii}
ggplot(gii_world) +
  geom_sf(aes(fill = value), color = "grey30", linewidth = 0.1) +
  scale_fill_distiller(
    palette = "YlOrRd", direction = 1,
    name = "GII", na.value = "grey80",
    limits = c(0, 1)
  ) +
  coord_sf(ylim = c(-56, 90)) +
  labs(title = "Gender Inequality Index, 2023") +
  theme(axis.text = element_blank(), axis.ticks = element_blank())
```

## Research Questions

1. Are there any gender-specific differences in:
   i. **education**,
   ii. **labor force participation**,
   iii. **care work**, and
   iv. **income**?
2. How do these gender-specific differences vary across countries, and how have they evolved over time?
3. Is there a correlation between these differences and development-related indicators such as the HDI, the Democracy Index, or others?

## Data sets

Our analysis is based on data sets gathered from the following sources:

-   ILOSTAT (<https://ilostat.ilo.org/>)
-   UNESCO Institute of Statistics (<https://databrowser.uis.unesco.org/>)
-   Our World in Data (<https://ourworldindata.org/>)

## Methodology

-   Single missing values were imputed using the average of the adjacent years

# Education

---

```{r gpi-all-combined, fig.width=12, fig.height=7}
(gpi_subplot("Primary") + gpi_subplot("Lower secondary")) /
(gpi_subplot("Upper secondary") + gpi_subplot("Tertiary")) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "Gender Parity Index by Education Level",
    subtitle = "Ratio of female to male gross enrollment ratio, 2020"
  ) &
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2, "cm")
  )
```

---

```{r completion-global-latest}
ggplot(completion.global.latest, aes(x = educationLevel, y = value, fill = sex)) +
  geom_bar(position = "dodge", stat = "identity", color = "black", linewidth = 0.5) +
  ylim(0, 100) +
  labs(title = "Average completion rates", subtitle = "Shown by Sex and Education level, 2021", x = "Education level", y = "Average completion rate", fill = "Sex")
```

---

```{r completion-countries}
ggplot(completion.countries, aes(x = year, y = value, color = sex)) +
  geom_line() +
  geom_point() +
  facet_wrap(~country, nrow = 2) +
  theme(panel.spacing.x = unit(2.5, "lines")) +
  labs(title = "Completion rates for selected countries over time", subtitle = "Averaged across primary, lower secondary, and upper secondary education", x = "Year", y = "Average completion rate", color = "Sex")
```

# Labor force participation

---

```{r}
lf_total_2024_comparison %>% 
  ggplot(aes(x = country, y = rate, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Labour Force Participation Rate in 2024 by Gender",
        x = "Country",
        y = "Labour Force Participation Rate in %",
        fill = "Biological Gender") +
  scale_fill_manual(values = c("Female" = "red", "Male" = "blue")) +
  scale_y_continuous(limits = c(0,90), breaks = seq(0,90,20)) +
  theme_light()
```

---

```{r}
lf_GER_2010_to_2024 %>%
  ggplot(aes(x = year, y = rate, colour = sex)) + geom_line() + geom_point() +
  labs(title = "Labour Force Participation Rate in Germany between 2010 and 2024",
       x = "Year",
       y = "Labour Force Participation Rate in %",
       colour = "Biological Gender") +
  scale_color_manual(values = c("Female" = "red", "Male" = "blue")) +
  scale_y_continuous(limits = c(0,70), breaks = seq(0,70,10)) + 
  theme_light()
```

---

```{r}
lf_GER_2010_to_2024 %>%
  ggplot(aes(x = year, y = rate, colour = sex)) + geom_line() + geom_point() +
  labs(title = "Labour Force Participation Rate in Germany between 2010 and 2024",
       subtitle = "Zoomed in",
       x = "Year",
       y = "Labour Force Participation Rate in %",
       colour = "Biological Gender") +
  scale_color_manual(values = c("Female" = "red", "Male" = "blue")) +
  theme_light() 
```

---

```{r}
lf_2024_by_education %>% filter(country == "Germany") %>%
  ggplot(aes(x = country, y = rate, fill = sex)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Labour Force Participation Rate by Gender and Education Level",
       subtitle = "Germany, 2024",
        x = "Country",
        y = "Labour Force Participation Rate in %",
        fill = "Biological Gender") +
  scale_fill_manual(values = c("Female" = "red", "Male" = "blue")) +
  facet_wrap(~ education,
             labeller = label_wrap_gen(width = 30)) +
  theme_light()
```

---

```{r}
lf_2024_by_education %>% filter(country == "Iran") %>%
  ggplot(aes(x = country, y = rate, fill = sex)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Labour Force Participation Rate by Gender and Education Level",
       subtitle = "Iran, 2024",
        x = "Country",
        y = "Labour Force Participation Rate in %",
        fill = "Biological Gender") +
  scale_fill_manual(values = c("Female" = "red", "Male" = "blue")) +
  facet_wrap(~ education,
             labeller = label_wrap_gen(width = 30)) +
  theme_light()
```

# Care Work

## Differences in Care Work

```{r}
plot_domestic_work_region(owid_domestic_work_time)
```

## Care Responsibilities Keeping People Out of Work

```{r}
plot_care_share_facet(care_responsibilities_share)
```

## Gap in Care Responsibility Share Amongst Countries

```{r}
plot_care_country_gap(care_responsibilities_share)
```

## Labour Force Participation vs Care Responsibility Share

```{r}
plot_care_lfp_correlation(care_responsibilities_share, lfp)
```

# Income
   
## Median wage gap over time
```{r}
plot_wage_gap_trend(trend_gap)
```

---

## Top countries with highest wage gap

```{r}
plot_wage_gap_top_countries(top_countries, latest_year)
```

---

## Largest changes over time

```{r}
plot_wage_gap_change(change_full, first_year, last_year)
```

---

## Change around the pandemic (2019–2020)

```{r}
plot_wage_gap_change(change_1920, 2019, 2020)
```

# Conclusion

<!-- TODO -->

## Summary

<!-- TODO -->

## Limitations

-   Missing values impede data aggregation on a global scale
-   Data shows limited availability across indicators examined

# Thank you very much!
