---
title: "Gender inequality"
subtitle: "Statistical project (LMU Munich)"
author: "Andreas Sünder, Miriam Haf, Jinwook Bae, Xuchen Shen"
date: today
date-format: "MMMM D, YYYY"
lang: en
execute:
  echo: false
format:
  revealjs:
    fontsize: 32px
    theme: default
    slide-number: true
    logo: assets/logo.svg
    css: presentation.css
---

```{r setup, include=FALSE}
# Put any `library()` calls (or similar) here
# We can't put those into the setup block below, as the functions loaded by them
# are not restored correctly when cached
source("source_all.R")
download_data()
```

```{r data-prep, include=FALSE, cache=TRUE, cache.extra=paste(file.info(c("source_all.R", "settings.R", list.files("code", pattern="\\.R$", full.names=TRUE)))$mtime, collapse="|")}
# You can put any constant values here
COUNTRIES_FOR_COMPLETION <- c("Germany", "Brazil", "Nigeria", "Dominican Republic")

world <- ne_countries(scale = "medium", returnclass = "sf")

gii <- load_owid_gii_data()
gii.latest <- gii[year == 2023]
gii_world <- merge(world, gii.latest, by.x = "iso_a3", by.y = "countryCode", all.x = TRUE)

completion <- load_uis_completion_rates()
balanced.countries.compl <- completion[year == 2021,
  if (uniqueN(educationLevel) == 3 && uniqueN(sex) == 2) .SD,
  by = country
]$country |> unique()
completion.global.latest <- completion[year == 2021 & country %in% balanced.countries.compl,
  .(value = mean(value)), by = .(educationLevel, sex)
]
completion.countries.time <- completion[between(year, 2005, 2020) & country %in% COUNTRIES_FOR_COMPLETION, .(value = mean(value)), by = .(country, year, sex)]
completion.countries.time[, country := factor(country, levels = COUNTRIES_FOR_COMPLETION)]

gpi <- load_owid_gpi_data()
# we outsource the repeating tasks for each education level
# to avoid duplicated code segments
plot_gpi_map <- function(data, level) {
  # Filter for 2020 and 2021, then pick the latest available year per country
  gpi_filtered <- data[year %in% c(2020, 2021) & educationLevel == level]
  gpi_filtered <- gpi_filtered[, .SD[which.max(year)], by = .(countryCode, country)]
  world_merged <- merge(world, gpi_filtered, by.x = "iso_a3", by.y = "countryCode", all.x = TRUE)

  ggplot(world_merged) +
    geom_sf(aes(fill = value), color = "grey30", linewidth = 0.1) +
    scale_fill_presentation_diverging(
      name = "GPI", na.value = "grey80",
      low = unname(config.palette.sex["Female"]),
      high = unname(config.palette.sex["Male"]),
      midpoint = 1,
      limits = c(0.5, 1.5), oob = scales::squish
    ) +
    coord_sf(ylim = c(-56, 90)) +
    labs(
      title = "Gender parity index",
      subtitle = "Latest available year in 2020-2021 | Female-to-male gross enrollment ratio"
    ) +
    theme(legend.title = element_text(vjust = 0.75), axis.text = element_blank(), axis.ticks = element_blank())
}

# this is an experimental function that is used by
# the combined GPI plot as it requires inducing some custom
# spacing to make the map use the available space efficiently
gpi_subplot <- function(level) {
  plot_gpi_map(gpi, level) +
    labs(title = level, subtitle = NULL) +
    theme(
      plot.title = element_text(size = 12, hjust = 0.5, margin = margin(b = 3)),
      plot.margin = margin(t = 8, r = 2, b = 2, l = 2)
    )
}

labour_force_with_educ <- load_labour_force_with_educ_data()
labour_force_with_educ_cleaned <- prepare_lfpr_data(labour_force_with_educ)

lf_total_2024_comparison <-
  prepare_lfpr_total_2024_comp(labour_force_with_educ_cleaned)

lf_GER_2010_to_2024 <-
  prepare_lfpr_GER_2010_to_2024(labour_force_with_educ_cleaned)

lf_2024_by_education <-
  prepare_lfpr_2024_by_education(labour_force_with_educ_cleaned)
lf_2024_educ_groupped <- group_educ_and_recalculate_rates(lf_2024_by_education)
lf_2024_tert_educ <- tertiary_educ_detailed(lf_2024_by_education)


wage_gap <- load_gender_wage_gap_data()

# Load care work and domestic work data
owid_domestic_work_time <- load_care_time_share_data()
care_responsibilities_share <- load_care_resp_share_data()
lfp <- load_lfp_data()
```

## Agenda

1. Motivation
2. Data Analysis:

    - Gender differences in **education**, **labor force participation**, **care work**, and **income**
    - Variation across countries and over time
    - Correlation with other development-related indicators
   
3. Summary
4. Discussion

# Motivation

---

```{r gii}
ggplot(gii_world) +
  geom_sf(aes(fill = value), color = "grey30", linewidth = 0.1) +
  scale_fill_presentation_sequential(
    name = "GII", na.value = "grey80",
    limits = c(0, 1)
  ) +
  coord_sf(ylim = c(-56, 90)) +
  labs(
    title = "Gender inequality index",
    subtitle = "Global coverage | 2023",
    caption = "Source: UNDP via Our World in Data.\nHigher values indicate greater inequality towards women."
  ) +
  theme(axis.text = element_blank(), axis.ticks = element_blank())
```

## Data sets & methodology

Our analysis is based on data sets gathered from the following sources:

-   ILOSTAT (<https://ilostat.ilo.org/>)
-   UNESCO Institute of Statistics (<https://databrowser.uis.unesco.org/>)
-   Our World in Data (<https://ourworldindata.org/>)

# Education

---

```{r gpi-all-combined, fig.width=12, fig.height=7}
(gpi_subplot("Primary") + gpi_subplot("Lower secondary")) /
(gpi_subplot("Upper secondary") + gpi_subplot("Tertiary")) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "Gender parity index",
    subtitle = "Global coverage | By education level | Latest available year in 2020-2021",
    caption = paste(
      "Source: Our World in Data.\n",
      "Values show the female-to-male gross enrollment ratio; 1 indicates parity."
    )
  ) &
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2, "cm")
  )
```

---

```{r completion-global-latest}
ggplot(completion.global.latest, aes(x = educationLevel, y = value, fill = sex)) +
  geom_bar(
    position = "dodge",
    stat = "identity",
    color = unname(config.palette.presentation$ink),
    linewidth = 0.5
  ) +
  ylim(0, 100) +
  scale_fill_sex() +
  labs(
    title = "Average completion rates",
    subtitle = "49 countries | By sex and education level | 2021",
    x = "Education level",
    y = "Average completion rate (%)",
    fill = "Sex",
    caption = paste(
      "Source: UNESCO Institute of Statistics.\n",
      "Includes countries with complete observations across all three education levels and both sexes in 2021."
    )
  )
```

---

```{r completion-countries-time}
ggplot(completion.countries.time, aes(x = year, y = value, color = sex)) +
  geom_line() +
  geom_point() +
  ylim(40, 100) +
  scale_color_sex() +
  facet_wrap(~country, nrow = 2) +
  theme(panel.spacing.x = unit(2.5, "lines")) +
  labs(
    title = "Completion rates",
    subtitle = "Selected countries | Averaged across education levels | 2005-2020",
    x = "Year",
    y = "Average completion rate (%)",
    color = "Sex",
    caption = paste(
      "Source: UNESCO Institute of Statistics.\n",
      "Values are country-year means across primary, lower secondary, and upper secondary completion rates."
    )
  )
```

# Labor force participation

---

```{r - Labor force participation, country comparison}
lf_total_2024_comparison %>% 
  ggplot(aes(x = country, y = rate, fill = sex)) +
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = unname(config.palette.presentation$ink),
    linewidth = 0.5
  ) +
  labs(title = "Labor force participation rates",
       subtitle = "Selected countries | By sex | 2024",
         x = "Country",
         y = "Labor force participation rate (%)",
         fill = "Sex",
         caption = "Source: ILOSTAT.") +
  scale_fill_sex() +
  scale_y_continuous(limits = c(0, 90), breaks = seq(0, 90, 20))
```

---

```{r - Labor force participation in Germany over time}
lf_GER_2010_to_2024 %>%
  ggplot(aes(x = year, y = rate, color = sex)) + geom_line() + geom_point() +
  labs(title = "Labor force participation rate",
       subtitle = "Germany | By sex | 2010-2024",
       x = "Year",
       y = "Labor force participation rate (%)",
       color = "Sex",
       caption = "Source: ILOSTAT.") +
  scale_color_sex() +
  scale_x_continuous(limits = c(2010, 2025), breaks = c(2010, 2015, 2020, 2025)) +
  scale_y_continuous(limits = c(0, 70), breaks = seq(0, 70, 10))
```

---

```{r - Labor force participation in Germany by education}
lf_2024_educ_groupped %>% filter(country == "Germany") %>%
  ggplot(aes(x = country, y = rate, fill = sex)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = unname(config.palette.presentation$ink),
    linewidth = 0.5
  ) + 
  labs(title = "Labor force participation rate",
       subtitle = "Germany | By sex and education level | 2024",
       x = NULL,
            y = "Labor force participation rate (%)",
            fill = "Sex",
            caption = "Source: ILOSTAT.") +
  scale_fill_sex() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  facet_wrap(~ education_new)
```

---

```{r - Labor force participation in Iran by education}
lf_2024_educ_groupped %>% filter(country == "Iran") %>%
  ggplot(aes(x = country, y = rate, fill = sex)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = unname(config.palette.presentation$ink),
    linewidth = 0.5
  ) + 
  labs(title = "Labor force participation rate",
       subtitle = "Iran | By sex and education level | 2024",
            x = NULL,
            y = "Labor force participation rate (%)",
            fill = "Sex",
            caption = "Source: ILOSTAT.") +
  scale_fill_sex() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  facet_wrap(~ education_new)
```

---

```{r - Labor force participation in Iran, tertiary education}
lf_2024_tert_educ %>% filter(country == "Iran") %>%
  ggplot(aes(x = country, y = rate, fill = sex)) +
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = unname(config.palette.presentation$ink),
    linewidth = 0.5
  ) +
  labs(title = "Labor force participation rate",
       subtitle = "Iran | By sex and tertiary education | 2024",
            x = NULL,
            y = "Labor force participation rate (%)",
            fill = "Sex",
            caption = "Source: ILOSTAT.\nRestricted to detailed tertiary education categories.") +
  scale_fill_sex() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  facet_wrap(~ education)
```


# Care work

---

```{r}
plot_domestic_work_region(owid_domestic_work_time)
```

---

```{r}
plot_care_share_facet(care_responsibilities_share)
```

---

```{r}
plot_care_country_gap(care_responsibilities_share)
```

---

```{r}
plot_care_lfp_correlation(care_responsibilities_share, lfp)
```

# Income
 
---

```{r}
plot_wage_gap_distribution(wage_gap)
```

---

```{r}
plot_wage_gap_facet(wage_gap)
```

---

```{r}
plot_wage_gap_gii_correlation(wage_gap)
```

<!-- Legacy income slides kept for reference.
Old slide 1: plot_wage_gap_trend(trend_gap)
Old slide 2: plot_wage_gap_top_countries(top_countries, latest_year)
Old slide 3: plot_wage_gap_change(change_full, first_year, last_year)
Old slide 4: plot_wage_gap_change(change_1920, 2019, 2020)
-->

# Conclusion

<!-- TODO -->

## Summary

<!-- TODO -->

## Discussion

-   Missing values impede data aggregation on a global scale
-   Data shows limited availability across indicators examined

# Thank you very much!
