---
title: "Gender inequality"
subtitle: "Statistical project (LMU Munich)"
author: "Andreas Sünder, Miriam Haf, Jinwook Bae, Xuchen Shen"
date: today
date-format: "MMMM D, YYYY"
lang: en
execute:
  echo: false
format:
  revealjs:
    fontsize: 32px
    theme: default
    slide-number: true
    logo: assets/logo.svg
    css: presentation.css
---

```{r setup, include=FALSE}
# Put any `library()` calls (or similar) here
# We can't put those into the setup block below, as the functions loaded by them
# are not restored correctly when cached
source("source_all.R")
download_data()
```

```{r data-prep, include=FALSE, cache=TRUE}
# You can put any constant values here
COUNTRIES_FOR_COMPLETION <- c("Germany", "Brazil", "Nigeria", "Dominican Republic")

world <- ne_countries(scale = "medium", returnclass = "sf")

gii <- load_owid_gii_data()
gii_world <- prepare_gii_world_map_data(gii, world)

gpi <- load_owid_gpi_data()

completion <- load_uis_completion_rates()
completion.global.latest <- prepare_completion_global_latest(completion)
completion.countries.time <- prepare_completion_countries_time(completion, COUNTRIES_FOR_COMPLETION)

lfp <- load_lfp_data()
labour_force_with_educ <- load_labour_force_with_educ_data()
labour_force_with_educ_cleaned <- prepare_lfpr_data(labour_force_with_educ)
lf_total_2024_comparison <-
  prepare_lfpr_total_2024_comp(labour_force_with_educ_cleaned)
lf_GER_2010_to_2024 <-
  prepare_lfpr_GER_2010_to_2024(labour_force_with_educ_cleaned)
lf_2024_by_education <-
  prepare_lfpr_2024_by_education(labour_force_with_educ_cleaned)
lf_2024_educ_groupped <- group_educ_and_recalculate_rates(lf_2024_by_education)
lf_2024_tert_educ <- tertiary_educ_detailed(lf_2024_by_education)

owid_domestic_work_time <- load_care_time_share_data()
care_responsibilities_share <- load_care_resp_share_data()

wage_gap <- load_gender_wage_gap_data()
```

## Agenda

1. Motivation
2. Data Analysis:

    - Gender differences in **education**, **labor force participation**, **care work**, and **income**
    - Variation across countries and over time
    - Correlation with other development-related indicators
   
3. Summary
4. Discussion

# Motivation

---

```{r gii}
plot_gii_map(gii_world)
```

## Data sets & methodology

Our analysis is based on data sets gathered from the following sources:

-   ILOSTAT (<https://ilostat.ilo.org/>)
-   UNESCO Institute of Statistics (<https://databrowser.uis.unesco.org/>)
-   Our World in Data (<https://ourworldindata.org/>)

# Education

---

```{r gpi-all-combined, fig.width=12, fig.height=7}
plot_gpi_combined(gpi, world)
```

---

```{r completion-global-latest}
plot_completion_global(completion.global.latest)
```

---

```{r completion-countries-time}
plot_completion_countries_time(completion.countries.time)
```

# Labor force participation

---

```{r - Labor force participation, country comparison}
plot_lfpr_selected_countries(lf_total_2024_comparison)
```

---

```{r - Labor force participation in Germany over time}
plot_lfpr_germany_time(lf_GER_2010_to_2024)
```

---

```{r - Labor force participation in Germany by education}
plot_lfpr_by_education(lf_2024_educ_groupped, "Germany")
```

---

```{r - Labor force participation in Iran by education}
plot_lfpr_by_education(lf_2024_educ_groupped, "Iran")
```

---

```{r - Labor force participation in Iran, tertiary education}
plot_lfpr_tertiary(lf_2024_tert_educ, "Iran")
```


# Care work

---

```{r}
plot_domestic_work_region(owid_domestic_work_time)
```

---

```{r}
plot_care_share_facet(care_responsibilities_share)
```

---

```{r}
plot_care_country_gap(care_responsibilities_share)
```

---

```{r}
plot_care_lfp_correlation(care_responsibilities_share, lfp)
```

# Income
 
---

```{r}
plot_wage_gap_distribution(wage_gap)
```

---

```{r}
plot_wage_gap_facet(wage_gap)
```

---

```{r}
plot_wage_gap_gii_correlation(wage_gap)
```

# Conclusion

<!-- TODO -->

## Summary

<!-- TODO -->

## Discussion

-   Missing values impede data aggregation on a global scale
-   Data shows limited availability across indicators examined

# Thank you very much!
